---
title: "p8105_hw1_ra3429"
author: "Robert Ariss"
date: "2026-09-26"
output: github_document
---

This is an R markdown document for homework 1

# Section 0: Libraries

```{r echo=TRUE}
library(tidyverse)
```

# Section 1: Data Import
```{r}
data("penguins", package = "palmerpenguins")
```

# Section 2: Description of the Penguins Data Set

The penguins data set consists of the following:

* There are `r nrow(penguins)` rows within the penguins data set. 

* There are `r ncol(penguins)` columns within the penguins data set. 

* The `r nrow(penguins)` rows represent data on each individual penguin

* The `r ncol(penguins)` columns represent characteristics of each penguin including `r colnames(penguins)` 

* For example, the mean flipper length in penguins without missing data was `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)`

Scatterplot
```{r echo=FALSE, warning=FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("Scatterplot.jpg")
```

# Section 3: Variable Types and Coercion

We will create a dataframe with multiple types of variables as below:
```{r}
problem_3_df = 
  tibble(
    problem_3_num = rnorm(10),
    problem_3_fac = problem_3_num > 0,
    problem_3_char = c("my", "name", "is", "rob", "this", "is", "the", "homework", "one", "submission" ),
    problem_3_factor = factor(c("NYHAI", "NYHAII", "NYHAIII", "NYHAI", "NYHAII", "NYHAIII", "NYHAI", "NYHAII", "NYHAIII", "NYHAI"))
  )
```

We will attempt to take the mean of each of these types of variables:
```{r}
mean(pull(problem_3_df, problem_3_num))
mean(pull(problem_3_df, problem_3_fac))
mean(pull(problem_3_df, problem_3_char))
mean(pull(problem_3_df, problem_3_factor))
```

We are unable to calculate a mean for character or factor type variables.

Instead, we will try to convert variables to numeric:

```{r echo = TRUE, results = 'hide'}
as.numeric(pull(problem_3_df, problem_3_fac))
as.numeric(pull(problem_3_df, problem_3_char))
as.numeric(pull(problem_3_df, problem_3_factor))
```

We are only able to convert the logical and factor variables and not the string type character variable to numeric. We cannot convert the character string type variable to a meaningful number, so we cannot take a mean of this variable.  




  
  


